SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2021-01-21 maintex.tex
spfit section rewritten

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of \textbf{169} star-forming galaxies in close pairs with those of mass-matched \textbf{control} galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.25 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy. We also find that the sSFR offsets in the outskirts of the paired galaxies are dependent on whether the galaxy is the more massive or less massive companion in the pair. The more massive companion experiences zero to a positive sSFR enhancement while the less massive companion experiences sSFR suppression in their outskirts. \textbf{Our results are consistent with the idea that merger induced gas-inflows are capable of triggering elevated star formation in paired galaxies. }
\end{abstract}

@@

\subsection{\textbf{Emission Line Extraction}}\label{sec:spfit}

\textbf{
We use our own {\sc idl} based spectral fitting code, {\sc spfit}\footnote{\url{https://github.com/fuhaiastro/spfit}}, to model the spectra from the MaNGA datacubes \citepalias{Fu:2018}. {\sc spfit} simultaneously fits emission lines and the stellar continuum with the Levenberg-Marquardt nonlinear least-squares minimization algorithm. Emission lines are parameterized as a Gauss-Hermite series and the stellar continuum is the sum of simple stellar populations (SSPs) convolved with the line-of-sight velocity distribution (LOSVD). 
}

\textbf{
{\sc SPFIT} re-bins the MaNGA datacubes into 1\arcsec $\times$ 1\arcsec\ pixels to remove the covariances between neighboring pixels. This is in contrast with previous spectral fitting codes designed for the MaNGA survey like the MaNGA {\sc DAP} \citep{Belfiore:2019} and {\sc PIPE3D} \citep{Sanchez:2016a, Sanchez:2016b} which uses the original 0.5\arcsec $\times$ 0.5\arcsec\ grid. Within each 1\arcsec$\times$1\arcsec\ pixel, we use {\sc spfit} to get measurements of emission line fluxes, equivalent widths, velocities, and velocity dispersions of 19 emission lines and measurements of stellar masses, ages, [Fe/H], and the kinematics of the stellar populations.
}

\textbf{
The code uses the SSP templates from the MUISCAT library \citep{Vazdekis:2012} which allows a spectral range of 3465$-$9469 \AA\ with a uniform spectral resolution of FWHM $=$ 2.5 \AA. This MIUSCAT library is the combination of the MILES \citep{Sanchez-Blazquez:2006}, CaT \citep{Cenarro:2001}, and Indo-U.S. stellar libraries \citep{Valdes:2004}. The MaNGA {\sc DAP} uses the MILES-HC (Hierarchically Clustered MILES templates) library \citep{Westfall:2019} and {\sc PIPE3D} library uses the GSD156 library \citep{Cid-Fernandes:2013} which is built from the GRANADA \citep{Martins:2005} and MILES libraries.
}

@@

\item At large radii (R/\reff\ $>$ 1.5), the more massive companion of pairs show a enhancement to the sSFR while the less massive companion of pairs show a suppression to the sSFR.

@@

%\appendix

>>>>

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of \textbf{169} star-forming galaxies in close pairs with those of mass-matched \textbf{control} galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.3 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy. We also find that the sSFR offsets in the outskirts of the paired galaxies are dependent on whether the galaxy is the more massive or less massive companion in the pair. The more massive companion experiences zero to a positive sSFR enhancement while the less massive companion experiences sSFR suppression in their outskirts. \textbf{Our results illustrate the complex tidal effects on star formation in closely paired galaxies. }
\end{abstract}

@@

\subsection{Spectral Fitting}\label{sec:spfit}

{\bf 
We use {\sc spfit} to model the MaNGA datacubes. The IDL package is publicly available\footnote{\url{https://github.com/fuhaiastro/spfit}} and was first used in our previous study of close galaxy pairs in MaNGA \citepalias{Fu:2018}. While data products from the MaNGA Data Analysis Pipeline \citep[DAP;][]{Belfiore:2019} and {\sc PIPE3D} \citep{Sanchez:2016a, Sanchez:2016b} are available, {\sc spfit} allows us to combine spectra inside a given aperture before fitting the combined spectrum. The feature alone made it better suited for this project. Additional features of {\sc spfit} include: (1) simultaneously fitting emission lines and stellar continuum, (2) coadding spaxels with either a Voronoi tessellation or arbitrary polygons while accounting for covariances, (3) modeling asymmetric emission lines, broad AGN emission lines, AGN continuum, and various dust-extinction laws, and (4) utilizing multi-threading when processing a large number of datacubes.

Here we provide a brief description of the fitting procedure of {\sc spfit}. Except for broad-line AGN, {\sc spfit} models the observed spectrum as a superposition of emission lines and simple stellar populations (SSPs). The SSP library of {\sc miuscat} \citep{Vazdekis:2012} is matched to the MaNGA spectral resolution and is convolved with the line-of-sight velocity distribution (LOSVD). Both the LOSVD and the profile of the emission lines are parameterized as separate Gauss-Hermite series \citep{van-der-Marel:1993} to the fourth order.

For model optimization, we prefer a fast algorithm like the Levenberg-Marquardt nonlinear least-squares minimization algorithm implemented in {\sc mpfit} \citep{Markwardt:2009}. But for complex spectral models like ours, the success of the fitting routine relies on a good initial ``guess'' solution, which can be provided by the penalized pixel-fitting method \citep[pPXF;][]{Cappellari:2017}. The pPXF method is robust because it solves the weights of the templates with a linear algorithm \citep{Lawson:1974} {\it independently} from solving the Gauss-Hermite LOSVD with a nonlinear optimizer ({\sc mpfit}). The {\sc spfit} package implements a three-stage fitting procedure: First, it masks out spectral regions around emission lines from the input spectrum and use pPXF on the masked spectrum with SSP-only templates to obtain the initial ``best-fit'' parameters of the stellar continuum; Then, it subtracts the best-fit stellar continuum model from the spectrum and use pPXF to fit the residual emission-line-only spectrum with Gaussian emission-line templates to obtain the initial ``best-fit'' parameters of the emission lines; Finally, it uses {\sc mpfit} on the full input spectrum with the two sets of ``best-fit'' parameters from pPXF to simultaneously fit all of the parameters describing the emission lines and the stellar continuum. Saved in the final FITS file are the best-fit parameters and their uncertainties, including emission-line fluxes, equivalent widths, and kinematics and luminosity-weighted stellar masses, ages, metallicities, and kinematics. 
}

@@

\item At large radii (R/\reff\ $>$ 1.5), the more massive companion of pairs show an enhancement to the sSFR while the less massive companion of pairs show a suppression to the sSFR.

@@
